---
title: "chap7.data_analysis"
author: "Hanh Nguyen"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### CHAPTER 7: EXPLORATORY DATA ANALYSIS ###

____________
INTRODUCTION
____________

Exploratory Data Analysis = EDA
```{r}
#load tidyverse
library(tidyverse)
```
- Goal:
    . To develop an understanding of your data by using questions as tools to guide your investigation so that focusing on a specific part of your dataset to decide which graphs, models, or transformations to make.
    . Ask quality questions to generate a large quantity of questions.
    
- Questions to look at:
    1. What type of variation occurs within my variables?
    2. What tyoe of covariation occurs between my variables?

- Definition:
    . Variable: a quantity, quality or property that you can measure
    . Value: the state of a variable when you measure it 
              --> Value may change from measurement to measurement
    . Observation: a set of measurements made under similar conditions.
                  An observations will contain several values, each associated with a different
                  variable.
                  ---> Observation can be referred as a data point.
    . Tabular data: a set of values, each associates with a variable and an observation.
            - Is *tidy* if each value is placed in its own "cell", each variable in its own column, and each observation in its own row

_________
Variation
_________
. Variation is the tendency of the values ofa variable to change from measurement to measurement.
        - It is true when measuring quantities, categoricals, or even across different subjects.
        - Every variable has its own pattern of variation

**Visualising distributions**

. Categorical variable : bar chart
                          - height : how many observations occurred with each x value
                          - compute using dplyr::count()
```{r}
ggplot(data=diamonds) + geom_bar(mapping=aes(x=cut))
diamonds %>% count(cut)
```

. Continuous variable  : histogram
  - date/time            - devides the x-axis into equally spaced bins, then, use the height
                            of a bar to display the nuber of observaions that fall in each
                            bin.
                          - if wish to overlay multiple histograms in the same plot
                            -> use geom-freqpoly() - same calculation but displaying the counts
                                with bars using lines instead
  - compute by combining dplyr::count() and ggplot2::cut_width()
```{r}
ggplot(data=diamonds) + geom_histogram(mapping=aes(x=carat), binwidth = 0.5)
diamonds %>% count(cut_width(carat, 0.5))
ggplot(data = diamonds, mapping = aes(x = carat, color = cut)) + geom_freqpoly(binwidth=0.1)
```
  
**Typical values**
. In both bar charts and histograms, tall bars show the common values of a variable, 
  and shorter bars show less common values. Places that do not have bars reveal values 
  that were not seen.
    - Useful questions: Most common? Rare? Why? Unusual patterns? Explain.
. Clusters of similar values suggest that subgroups exist.
    - Useful questions: Similliar? Different? Explain & describe. Misleading?
    
**Unusual values**
. Outliers are observations that are unusual, data points that don't seem to fit the pattern.
    - Either wrong dat entry or important new science.
    - coord_cartesian() : make it easy to see to unusual values
            ~ xlim() : zoom in x axis
            ~ ylim() : zoom in y axis
```{r}
# regular graph
ggplot(diamonds) + geom_histogram(aes(x = y), binwidth = 0.5)
# zoom in graph with coord_cartersian()
ggplot(diamonds) + 
  geom_histogram(aes(x = y), binwidth = 0.5) + 
  coord_cartesian(ylim = c(0, 50))
```



________________________
EXERCISE 7.3 - VARIATION
________________________
###1.    Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
diamonds
# transform diamonds from wide format to long format
diamonds_long <- gather(diamonds, side, measurment, x:z, factor_key = TRUE)
diamonds_long
# plot diamonds in long format
ggplot(diamonds_long, aes(x = measurment)) + geom_histogram(aes(fill=side))
```
--> x is length, y is width, z is depth.

###2.    Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)
```{r}
#Distribution of the price based on the color of the diamond
ggplot(diamonds, aes(x = price, color = color)) + geom_freqpoly()
# Distribution of the price based on the cut of the diamond
ggplot(diamonds, aes(x = price, color = cut)) + geom_freqpoly()
# Distribution of price in general in the range from $1000 to $2000
ggplot(diamonds, aes(x = price)) + geom_freqpoly() + xlim(c(1000, 2000))
```
--> It seems that there is no diamond which has price around $1500.

###3.    How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?
```{r}
# Count diamonds in 0.99 carat.
diamonds %>% filter(carat == 0.99) %>% count()
# Count diamonds in 1 carat.
diamonds %>% filter(carat == 1) %>% count()
# plot
ggplot(diamonds, aes(x = carat)) + geom_histogram() + xlim(c(0.97, 1.05))
```
--> There is much more 1 carat than 0.99 carat. It's probably because most people prefer buying their diamond in whole countable units.

###4.    Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?
```{r}
# xlim
ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = 0.01) + xlim(c(0.97, 1.05))
# coor_cartesian()
ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = 0.01) + coord_cartesian(xlim = c(0.97, 1.05))
```
--> xlim select the range first and plot the value in the range, thus the plot doesn't look continuous outside of the range
--> with coord_cartesian, the plot looks more continous to values outside of the range.

