---
title: "chap7.data_analysis"
author: "Hanh Nguyen"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### CHAPTER 7: EXPLORATORY DATA ANALYSIS ###


(EXERCISE STARTING AT LINE #270)


____________
INTRODUCTION
____________

Exploratory Data Analysis = EDA
```{r}
#load tidyverse
library(tidyverse)
```
- Goal:
    . To develop an understanding of your data by using questions as tools to guide your investigation so that focusing on a specific part of your dataset to decide which graphs, models, or transformations to make.
    . Ask quality questions to generate a large quantity of questions.
    
- Questions to look at:
    1. What type of variation occurs within my variables?
    2. What tyoe of covariation occurs between my variables?

- Definition:
    . Variable: a quantity, quality or property that you can measure
    . Value: the state of a variable when you measure it 
              --> Value may change from measurement to measurement
    . Observation: a set of measurements made under similar conditions.
                  An observations will contain several values, each associated with a different
                  variable.
                  ---> Observation can be referred as a data point.
    . Tabular data: a set of values, each associates with a variable and an observation.
            - Is *tidy* if each value is placed in its own "cell", each variable in its own column, and each observation in its own row

_________
Variation
_________
. Variation is the tendency of the values ofa variable to change from measurement to measurement.
        - It is true when measuring quantities, categoricals, or even across different subjects.
        - Every variable has its own pattern of variation

**Visualising distributions**

. Categorical variable : bar chart
                          - height : how many observations occurred with each x value
                          - compute using dplyr::count()
```{r}
ggplot(data=diamonds) + geom_bar(mapping=aes(x=cut))
diamonds %>% count(cut)
```

. Continuous variable  : histogram
  - date/time            - devides the x-axis into equally spaced bins, then, use the height
                            of a bar to display the nuber of observaions that fall in each
                            bin.
                          - if wish to overlay multiple histograms in the same plot
                            -> use geom-freqpoly() - same calculation but displaying the counts
                                with bars using lines instead
  - compute by combining dplyr::count() and ggplot2::cut_width()
```{r}
ggplot(data=diamonds) + geom_histogram(mapping=aes(x=carat), binwidth = 0.5)
diamonds %>% count(cut_width(carat, 0.5))
ggplot(data = diamonds, mapping = aes(x = carat, color = cut)) + geom_freqpoly(binwidth=0.1)
```
  
**Typical values**
. In both bar charts and histograms, tall bars show the common values of a variable, 
  and shorter bars show less common values. Places that do not have bars reveal values 
  that were not seen.
    - Useful questions: Most common? Rare? Why? Unusual patterns? Explain.
. Clusters of similar values suggest that subgroups exist.
    - Useful questions: Similliar? Different? Explain & describe. Misleading?
    
**Unusual values**
. Outliers are observations that are unusual, data points that don't seem to fit the pattern.
    - Either wrong dat entry or important new science.
    - coord_cartesian() : make it easy to see to unusual values
            ~ xlim() : zoom in x axis
            ~ ylim() : zoom in y axis
```{r}
# regular graph
ggplot(diamonds) + geom_histogram(aes(x = y), binwidth = 0.5)
# zoom in graph with coord_cartersian()
ggplot(diamonds) + 
  geom_histogram(aes(x = y), binwidth = 0.5) + 
  coord_cartesian(ylim = c(0, 50))
```

______________
Missing Values
______________
. Encountering unusual values:
    1. Drop the entire row with the strange values
        -> not recommended
    2. Replacing the unusual values with missing values.
        -> recommended
          - *mutate()* to replace the variable with a modified copy
          - *ifelse()* to replace unusual values with NA
                    ~ test : a logical vector
                    ~ yes : result of test if test is TRUE
                    ~ no : result of test if test is FALSE
                Alternative: *dplyr::case_when()* - creates a new
                                    variable that relies on a complex
                                    combination of existing variables.

```{r}
diamonds
# option1: drop entire row
diamonds2 <- diamonds %>% 
  filter(between(y, 3, 20))
diamonds2
# option2: replacing with NA
diamonds3 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))
# plot with missing value
ggplot(data = diamonds3, mapping = aes(x = x, y = y)) + 
  geom_point()
# Suppress warning as missing values are removed, set na.rm = TRUE.
ggplot(data = diamonds3, mapping = aes(x = x, y = y)) + 
  geom_point(na.rm = TRUE)
```

___________
Covariation
___________
. Variation describes the behavior within a variable.
. Covariation describes the behavior between variables.
. Covariation is the tendency for the values of two or more variables
  to vary together in a related way.

**A categorical and continuous variable**
. Display *density*, which is the count standardized so that the area
  under each frequency polygon is one -> make comparison easier.
```{r}
ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) +
  geom_freqpoly(mapping = aes(color = cut), binwidth = 500)
```
. *Boxplot* -  a type of visual shorthand for a distribution of
              values, consisting:
                ~ A box that stretches from the 25th percentile of the
                  distribution to the 75th percentile, a distance 
                  known as the interquartile range (IQR). In the 
                  middle of the box is a line that displays the 
                  median, ie 50th percentile, of the distribution.
                  --> the spread of the distribution, whether or not
                      the distribution is symmetric about the median
                      or skewed to one side.
                ~ Visual points that display observation that fall
                  more than 1.5 times the IQR from either edge of the
                  box. These outlying point are unusual so are plotted
                  individually
                ~ A line (or whisker) that extends from each end of
                  the box and goes to the farthest non-outlier point
                  in the distribution
```{r}
ggplot(diamonds, aes(x=cut, y=price)) + geom_boxplot()
```
. *coord_flip()* - flip the coordinate 90degree

**Two categorical variables**
- To visualize the covariation between categorical variables, 
  you'll need to count the number of observations for each
  combination.
```{r}
# The size of the dot displays how many observations occurred at each combination of values
ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))
# Compute the count with dplyr
diamonds %>% 
  count(color, cut) %>% 
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n))
```

**Two continuous variables**
. Scatterplot is one way to visualized the covariation between 2 continuous variables
```{r}
#exponential relationship between the carat size and price of a diamond.
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))
```
. Scatterplot becomes less useful as the size of your dataset grows because points begin to overplot, and pile up into uniform black
    ---> *alpha()*
    ---> *geom_bin2d()* and *geom_hex()* - divide the coordinate plan into 2d bins and then
                                          use a fill color to display how many points fall
                                          into each bin.
                                    ~ *geom_bin2d()* creates rectangular bins.
                                    ~ *geom_hex()* creates hexagonal bins.
.
```{r}
# fix the prolem with large dataset by transparency in *alpha()*
ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price), alpha = 1 / 100)
# fix the prolem with large dataset by geom_bin2d()
smaller <- diamonds %>% filter(carat < 3)
ggplot(data = smaller) +
  geom_bin2d(mapping = aes(x = carat, y = price))
# fix the prolem with large dataset by geom_hex()
#install.packages("hexbin")
ggplot(data = smaller) +
  geom_hex(mapping = aes(x = carat, y = price))
```
. Another option is to bin one continuous variable so it acts like a categorical variable
  by geom_boxplot()
      ~ cut_width(x, width) - divides x into bins of width 
      ~ varwidth = TRUE  - make the width of the boxplot proportional to the # of point
      ~ cut_number() - display approximately the same # of points in each bin
```{r}
ggplot(data = smaller, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)))

ggplot(data = smaller, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```

___________________
Patterns and models
___________________
. Patterns 
  - provide clues about relationships between variables.
  - reveal covariation (variation creates uncertainty, covariation reduce uncertainty)
          ~ If two variable covary -> prediction of one variable from the other variable.
          ~ If the covariation is due to a causal relationship (a special case) -> use
            the value of one variable to control the value of the second
  - Questions:
            Could this pattern be due to coincidence (ie, randome chance)?
            How can you describe the relationship implied by the pattern?
            How strong is the relationship implied by the pattern?
            What other variables might affect the relationship?
            Does the relationship change if you look at individual subgroups of the data?
            
. Models 
  - a tool for extracting patterns out of data.
  - It's possible to use a model to remove the very strong relationship between variables
```{r}
library(modelr)

mod <- lm(log(price) ~ log(carat), data = diamonds)

diamonds2 <- diamonds %>% 
  add_residuals(mod) %>% 
  mutate(resid = exp(resid))

ggplot(data = diamonds2) + 
  geom_point(mapping = aes(x = carat, y = resid))
```

_________________
7.7 ggplot2 calls
_________________
. Formula:
      ggplot(data=, aes(x=, y=)) + geom_*()





________________
### EXERCISE ###
________________

________________________
EXERCISE 7.3 - VARIATION
________________________
###1.    Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.
```{r}
diamonds
# transform diamonds from wide format to long format
diamonds_long <- gather(diamonds, side, measurment, x:z, factor_key = TRUE)
diamonds_long
# plot diamonds in long format
ggplot(diamonds_long, aes(x = measurment)) + geom_histogram(aes(fill=side))
```
--> x is length, y is width, z is depth.

###2.    Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)
```{r}
#Distribution of the price based on the color of the diamond
ggplot(diamonds, aes(x = price, color = color)) + geom_freqpoly()
# Distribution of the price based on the cut of the diamond
ggplot(diamonds, aes(x = price, color = cut)) + geom_freqpoly()
# Distribution of price in general in the range from $1000 to $2000
ggplot(diamonds, aes(x = price)) + geom_freqpoly() + xlim(c(1000, 2000))
```
--> It seems that there is no diamond which has price around $1500.

###3.    How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?
```{r}
# Count diamonds in 0.99 carat.
diamonds %>% filter(carat == 0.99) %>% count()
# Count diamonds in 1 carat.
diamonds %>% filter(carat == 1) %>% count()
# plot
ggplot(diamonds, aes(x = carat)) + geom_histogram() + xlim(c(0.97, 1.05))
```
--> There is much more 1 carat than 0.99 carat. It's probably because most people prefer buying their diamond in whole countable units.

###4.    Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?
```{r}
# xlim
ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = 0.01) + xlim(c(0.97, 1.05))
# coor_cartesian()
ggplot(diamonds, aes(x = carat)) + geom_histogram(binwidth = 0.01) + coord_cartesian(xlim = c(0.97, 1.05))
```
--> xlim select the range first and plot the value in the range, thus the plot doesn't look continuous outside of the range
--> with coord_cartesian, the plot looks more continous to values outside of the range.

_______________________________
EXERCISE 7.4 - Missing Values
_______________________________
    
###1.    What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?
---> Historgram removes missing value on plot.
---> Bar chart contain missing value on plot as another catergory.

###2.    What does na.rm = TRUE do in mean() and sum()?
---> na.rm = TRUE ignore NA when calculating mean and sum

______________________________________________________
EXERCISE 7.5.1 - A categorical and continuous variable
______________________________________________________
###1.    Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.
```{r}
flights %>% 
  mutate(canceled_flight = is.na(dep_time) | is.na(arr_time)) %>%
  ggplot() + geom_boxplot(aes(x=canceled_flight, y=dep_time))
```

###2.    What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?
```{r}
diamonds
# Carat and cut
ggplot(diamonds, aes(x= cut, y=carat)) + geom_boxplot()
# Carat, price and cut
ggplot(diamonds, aes(x=carat, y=price, color=cut)) + geom_point()
```
--> Carat is the most important variable to predict diamond price
    because more lower quality diamonds are produced in heavier carat.

###3.    Install the ggstance package, and create a horizontal boxplot. How does this compare to using coord_flip()?
```{r}
install.packages("ggstance")
library(ggstance)
ggplot(diamonds, aes(x= cut, y=carat)) + 
  geom_boxplot() + 
  coord_flip()
ggplot(diamonds, aes(x= carat, y=cut)) + 
  geom_boxploth()
```
--> They both give the same return

###4.    One problem with boxplots is that they were developed in an era of much smaller datasets and tend to display a prohibitively large number of “outlying values”. One approach to remedy this problem is the letter value plot. Install the lvplot package, and try using geom_lv() to display the distribution of price vs cut. What do you learn? How do you interpret the plots?
```{r}
#install.packages("lvplot")
library(lvplot)
ggplot(diamonds, aes(x=cut, y=price)) + geom_lv()
```
--> The letter value plot shows more quantile compared to boxplot

###5.    Compare and contrast geom_violin() with a facetted geom_histogram(), or a coloured geom_freqpoly(). What are the pros and cons of each method?
```{r}
ggplot(diamonds, aes(x=cut, y=price)) + geom_violin()
ggplot(diamonds, aes(x=price)) + geom_histogram() + facet_wrap(~cut)
ggplot(diamonds, aes(x=price, color=cut)) + geom_freqpoly() 
```
--> geom_violin give the best display for density while geom_freqpoly() can illustrate the price distribution in different cuts at the same time.On the other hand, geom_histogram() more focus on price distribution

###6.    If you have a small dataset, it’s sometimes useful to use geom_jitter() to see the relationship between a continuous and categorical variable. The ggbeeswarm package provides a number of methods similar to geom_jitter(). List them and briefly describe what each one does.
```{r}
#install.packages("ggbeeswarm")
library(ggbeeswarm)
ggplot(diamonds, aes(x=cut, y=price)) + 
  geom_jitter(size=0.1, color="pink")
ggplot(diamonds, aes(x=cut, y=price)) + geom_beeswarm()
```
-->`geom_quasirandom()` produces plots that are a mix of jitter 
                        and violin plots. There are several different
                        methods that determine exactly how the random
                        location of the points is generated.
-->`geom_beeswarm()` produces a plot similar to a violin plot, but 
                    by offsetting the points.


______________
EXERCISE 7.5.2
______________
###1.    How could you rescale the count dataset above to more clearly show the distribution of cut within colour, or colour within cut?
```{r}
# original
diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n))
# rescale
diamonds %>%
  count(color, cut) %>%
  group_by(color) %>%
  mutate(prop = n / sum(n)) %>%
  ggplot(mapping = aes(x = color, y = cut)) +
  geom_tile(mapping = aes(fill = prop))
```


###2.    Use geom_tile() together with dplyr to explore how average flight delays vary by destination and month of year. What makes the plot difficult to read? How could you improve it?

```{r}
# first glance plot
flights %>%
  group_by(month, dest) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(month), y = dest, fill = dep_delay)) +
  geom_tile()

# improved plot
flights %>%
  group_by(month, dest) %>%
  summarise(dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(x = factor(month), y = dest, fill = dep_delay)) +
  geom_tile() +
  scale_fill_viridis_b()
```


###3.    Why is it slightly better to use aes(x = color, y = cut) rather than aes(x = cut, y = color) in the example above?
```{r}
diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = cut, y = color)) +
    geom_tile(mapping = aes(fill = n))
```
--> Because it is easier to read with a horizontal catergorical variable.

______________
EXERCISE 7.5.3
______________
###1. Instead of summarising the conditional distribution with a boxplot, you could use a frequency polygon. What do you need to consider when using cut_width() vs cut_number()? How does that impact a visualisation of the 2d distribution of carat and price?

--> cut_width(x, width) divides x into bins of width --> width is concerned
--> cut_number() display approximately the same number of points in each bin -> # of bin

--> Bin widths and number need to be large enough so that it can remove noise, but not all
    the signal

###2. Visualise the distribution of carat, partitioned by price.
```{r}
# 15 bins
ggplot(diamonds, aes(x = cut_number(price, 15), y = carat)) +
  geom_boxplot() +
  coord_flip() 
#width of $1500, bodary = 0 ensure that the first bin is 0-1500
ggplot(diamonds, aes(x = cut_width(price, 1500, boundary = 0), y = carat)) +
  geom_boxplot(varwidth = TRUE) +
  coord_flip()
```

###3. How does the price distribution of very large diamonds compare to small diamonds? Is it as you expect, or does it surprise you?
--> Distribution of very large diamonds is more variable and more profitable.
--> Smaller diamonds are only profitable if they are more excetional in color and cut.

###4. Combine two of the techniques you’ve learned to visualise the combined distribution of cut, carat, and price.
```{r}
ggplot(diamonds, aes(x = cut_number(carat, 5), y = price, colour = cut)) +
  geom_boxplot()
```

###5. Two dimensional plots reveal outliers that are not visible in one dimensional plots. For example, some points in the plot below have an unusual combination of x and y values, which makes the points outliers even though their x and y values appear normal when examined separately.
```{r}
# original plot
ggplot(data = diamonds) +
  geom_point(mapping = aes(x = x, y = y)) +
  coord_cartesian(xlim = c(4, 11), ylim = c(4, 11))

```
Why is a scatterplot a better display than a binned plot for this case?
--> Because they have a very strong linear relationship, thus, better in scatterplot.
