---
title: "Titanic"
author: "Hanh Nguyen"
date: "4/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load Titanic
```{r}
titanic_raw <- read.csv("C:/Users/Jo/Desktop/CodingCult/practice-repo/titanic.csv")
titanic_raw
titanic <- titanic_raw %>% rename(Gender = Sex, Class = Pclass)
titanic
```
The dataset has 10 variables and 887 observations, each observation represents a passenger on the Titanic.
The 10 variables:
- Survived: 0 for not survived, 1 for survived
- Class: 1st, 2nd, and 3rd
- Name: passenger name
- Sex: male and female
- Age
- Siblings.Spouses.Aboard: number of siblings & spouses of the passenger who were also on board
- Parents.Children.Aboard: number of parents & children of the passenger who were also on board
- Fare: The cost of the fare that the passenger paid

_________________
GENDER & SURVIVAL
_________________
Purpose: to see if there is a relationship between gender and survival.

```{r}
#number of female passenger
total_female <- titanic %>% summarise(sum(Gender == "female"))
total_female
#number of male passeger
total_male <- titanic %>% summarise(sum(Gender == "male"))
total_male
#number of survived female passenger
survived_female <- titanic %>% count(Gender == "female" & Survived == 1)
survived_female
#number of unsurvived female passenger
unsurvived_female <- titanic %>% count(Gender == "female" & Survived == 0)
unsurvived_female
#number of survived male passenger
survived_male <- titanic %>% count(Gender == "male" & Survived == 1)
survived_male
#number of unsurvived male passenger
unsurvived_male <- titanic %>% count(Gender == "male" & Survived == 0)
unsurvived_male
#gender and survival plot
titanic %>%
  ggplot(aes(x = Gender, y = Survived)) + geom_col(width = 0.2) +
  labs(title = "Gender and Survival on the Titanic")
```
*Conclusion*
Female passengers had a significant higher chance of survival than male passengers.
This makes sense because during the Titanic crash, women were prioritized to get on the rescue boat.

______________
AGE & SURVIVAL
______________
Purpose: to see if there is a relationship between the passenger's age and their chance of survival.

```{r}
#age-survival
titanic %>%
  ggplot(aes(x = Age)) + geom_histogram() + facet_wrap(~Survived)

```
*Conclusion*
A large number of people were unable to survive the crash.
The age that had the highest number of death were 20 years.


________________
CLASS & SURVIVAL
________________
